---
title: "MERS Proteomics"
output: html_notebook
---

### clean environment
```{r}
rm(list=ls())
```

### set directory
```{r setup}
    knitr::opts_knit$set(root.dir = normalizePath("~/Desktop/Code-PHD/Proteomics_virus/")) 
```

```{r}
library(ggplot2)
library(limma)
library(dplyr)
library(xlsx)
library(RColorBrewer)
library(gplots)
library(ggrepel)
library(knitr)
library(ggfortify)
library(NormalyzerDE)
library(ropls)
library(stringr)
opts_chunk$set(echo = TRUE)
options(bitmapType="cairo")
```

```{r}
source("src/usefull_functions.R")
```

```{r}
name_ana <- "SARS"
input <- "data/SARS/20200818_L1_nUT2_SARS_TMT_Second_experiment.csv"

c_c1 <- "#83adb5"
c_c2 <- "#c7bbc9"
c_c3 <- "#8e768a"
c_c4 <- "#6c797c"
```

```{r}
dir.create(paste0("processing/", name_ana))
```

```{r}
dir.create("processing/normalization")
dir.create(paste0("results/", name_ana))
dir.create(paste0("results/", name_ana, "/figures"))
dir.create(paste0("results/", name_ana, "/DGE"))
dir.create(paste0("results/", name_ana, "/GSEA"))
```

## 1) Pre-processing

```{r}
data <- read.csv(input)
data$Description <- NULL
samples <- data.frame(names = colnames(data)[-1], condition = NA, names_2 = NA)

samples$disease <- ifelse(str_detect(samples$names, "SARS"), "SARS","UNINF")
samples$time <- ifelse(str_detect(samples$names, "24"), "24","48")
samples$condition <- paste0(substr(samples$disease, start = 1 , stop = 2), samples$time)

samples$names_2 <- paste0(samples$condition, "_", substr(gsub(".*\\.R", "",samples$names), start = 1, stop = 1)) ## extract run number
samples

condition <- samples$condition

colnames(data)[-1] <- samples$names_2
rownames(data) <- data$Accession
data$Accession <- NULL
```

## Make annotation_info_file
```{r}
annotationInfo <- read.delim("data/annotationInfo_Uniprot_gene_name")
names(annotationInfo)[1] <- "Accession"
```

## select Abundances columns
```{r}
print(nrow(data))
data <- data[rowSums(is.na(data)) != ncol(data), ] ## remove empty lines
print(nrow(data))
```
## save files
```{r}
save_file_csv(data, "row_data_filt", paste0("processing/", name_ana, "/"))

save_file_csv(samples, "informations", paste0("processing/", name_ana, "/"))
```

### check for missing values
```{r}
counts <- 1:nrow(data) # create vector of appropriate length
for(i in 1:nrow(data)){
    # TRUE will be coerced to 1 for the summing
    counts[i] <- sum(is.na(data[i,]))
}
table(counts) # create the summary
```
### distribution
```{r}
data_2 <-as.matrix(data)
hist(data_2 , prob=T)
curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
save_figure(data_2, "distribution_", 10, 10)
```


### PCA before norm
```{r}
pca_data <- data.frame(group = condition, 
                       t(as.matrix(data)))
```

```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-1])
```

```{r}
a <- autoplot(pca, data = pca_data, size= 6, shape = 21, fill = 'group')+  scale_fill_manual(values = c(c_c2, c_c3,c_c1, c_c4))+theme_gray()
print(a)
save_figure(a, "pca_before_norm_", 4, 4)
```


### Quantile normalization

```{r}
path_data <- make_data_matrix(data, name_ana)
path_design <-make_design_matrix(data, name_ana, condition)
outDir <- "processing/normalization"
```

```{r}
#normalyzer(jobName=name_ana, designPath=path_design, dataPath=path_data, outputDir=outDir)
```
```{r}
path_results_norm <- paste0(outDir, "/", name_ana,"/Quantile-normalized.txt")
data_norm <- read.delim(path_results_norm)
rownames(data_norm) <- rownames(data)

path_results <- paste0("processing/norm_files/", name_ana, "_norm_file.txt")

write.table(data_norm, file = path_results, 
            sep = "\t", row.names = TRUE, col.names = NA)
```
processing/normalization/SARS/Quantile-normalized.txt

### PCA after normalization
```{r}
pca_data <- data.frame(group = condition, t(as.matrix(data_norm)))
```

```{r}
pca_data[is.na(pca_data)] <-  0
pca <- prcomp(pca_data[,-1])
```

```{r}
b <- autoplot(pca, data = pca_data, size= 6, shape = 21, fill = 'group')+ scale_fill_manual(values = c(c_c2, c_c3,c_c1, c_c4))+theme_gray()
print(b)
save_figure(b, "pca_after_norm_", 4, 4)
```

```{r}
c <- autoplot(pca, data = pca_data, size= 6, fill = 'group', label = TRUE, label.size = 3, shape = FALSE)+ scale_fill_manual(values = c(c_c2, c_c3,c_c1, c_c4))+theme_gray()
save_figure(c, "pca_after_norm_label", 4, 4)
```

### distribution after normalization
```{r}
data_2 <-as.matrix(data_norm)
hist(data_2 , prob=T)
curve(dnorm(x,mean(data_2),sd(data_2)),add=T,col="red")
save_figure(data_2, "distribution_", 10, 10)
```
## Multiple comparisons

## four time points experiments = 

## set levels
```{r}
groups <- as.factor(condition)
design <- model.matrix( ~ 0 + groups)
colnames(design) <- levels(groups)
```

## fit model
```{r}
fit <- lmFit(data_norm, design)
```

```{r}
list_comp <- c("SA48-UN48", "SA24-UN24", "SA48-SA24", "UN48-UN24")
```

```{r}
head(data_norm)
```

```{r}
for(comp in list_comp){
  c2 <- substr(comp, 1,4)[1]
  c1 <- substr(comp, 6,9)[1]
  dge <- extract_limma(fit, design, name_ana, c1, c2)
  print(dge)
  dat <- select(data_norm, contains(c1), contains(c2))
  clustering_DGE(dat, dge, name_ana, comp, annotationInfo)
  top_table_up <- dge[dge$Significance_2 == "upregulated",]
  top_table_down <- dge[dge$Significance_2 == "downregulated",]
  print(top_table_up)
  name_up <- paste0("results/",name_ana,"/GSEA/prot_up_regulated_", comp, ".txt")
  name_down <- paste0("results/",name_ana,"/GSEA/prot_down_regulated_", comp, ".txt")
  print(name_up)
  print(name_down)
  write.table(data.frame(top_table_up$Gene_name), file = name_up, sep = "\t",
            row.names = FALSE)
  write.table(data.frame(data = top_table_down$Gene_name), file = name_down, sep = "\t",
            row.names = FALSE)
}
```

